# IRIS

## Species 분류

-   총 5개의 컬럼으로 구성

    ```{r}
    head(iris)
    ```

```{r}
plot(iris)
```

\-\-\-\-\-\-\-\-\-\-\-\--

iris의 첫 6개의 데이터

```{r}
head(iris)
```

iris의 마지막 6개의 데이터

```{r}
tail(iris)
```

iris의 데이터 프레임 구조

```{r}
str(iris)
```

iris의 데이터 프레임에 대한 요약

```{r}
summary(iris)
```

```{r}
plot(iris$Petal.Length)
```

# Vector

-   c(0), seq(), rep(), :으로 생성

# Factor(character, numeric)

```{r}
as.character(head(iris$Species))
as.numeric(c(head(iris$Species),tail(iris$Species)))
```

## Vector 계산

```{r}
x=c(3,6,9,12,15)
x+1 #x의 모든 값에 1 더하기
x+c(1,2,3,4,5)
```

## Vector 합치기

```{r}
x=c(3,5,6,12,15)
y=1:5
z=append(x,y);z #x에 y 덧붙이기
```

```{r}
z=c(x,y);z #누적 반복문에는 사용 불가
```

```{r}
length(x)==length(c(1,2))
x+c(1,2) #Not recommended
```

### iris의 Sepal.Length의 중복되지 않은 개수의 Percent 구하기

```{r}
length(unique(iris$Sepal.Length))/length(iris$Sepal.Length)*100
```

## Function related to String

```{r}
x=c("a","b","c","d")
x[1:2]
paste(x[1],x[2]) #Merge string
```

```{r}
paste("Hello", "World")
```

### Print 1st, 51st, 101st Species of iris

```{r}
paste(iris$Species[1], iris$Species[51], iris$Species[101])
```

```{r}
print("aaa") #Only one chunk can be printed
cat("aaa \t bbbb \n ccccc \rddd")
```

```{r}
paste("010","1111","2222", sep="-") #There is a blank between datas in default
```

### Extract String

```{r}
str01="abcdefghijklm"
substring(str01,2,5) #Extract String from str01 index from 2 to 5
```

### Check if data exists

```{r}
x=runif(5,1,20);x
```

```{r}
sum((x>=4)&(x<=7))
```

```{r}
any(x>=10) #Check if any of x is over 10
all(x>=10) #Check if all of x is over 10
```

# Matrix

-   Matrix contains 2 dimension of data

-   All types of data must be same to make matrix

```{r}
vec1=c(1,2,3)
vec2=c(4,5,6)
vec3=c(7,8,9)
```

rbind: Bind row

```{r}
mat1=rbind(vec1, vec2, vec3);mat1
```

cbind: Bind column

```{r}
mat2=cbind(vec1,vec2,vec3);mat2
```

```{r}
irisHead=head(iris,5);irisHead
```

```{r}
irisHead=rbind(irisHead,c(1,1,1,1,"setosa"));irisHead
```

```{r}
irisHead=cbind(irisHead, abc=rep(2,6));irisHead
```

```{r}
irisHead=iris[1:5,]
irisHead=cbind(irisHead, Length=(irisHead$Sepal.Length+irisHead$Petal.Length));irisHead
```

```{r}
irisHead=iris[1:5,]
irisHead$Length=irisHead$Sepal.Length+irisHead$Petal.Length
irisHead
```

```{r}
sl=iris$Sepal.Length
sw=iris$Sepal.Width
irisM=cbind(sl, sw)
head(irisM)
class(irisM)
```

```{r}
irisD=as.data.frame(irisM);head(irisD)
class(irisD)
```

Naming column of matrix

```{r}
colnames(mat1)=c("A", "B", "C")
mat1
```

Create mat1D as data frame of mat1

```{r}
mat1D=as.data.frame(mat1);mat1D
```

Naming row of matrix

```{r}
rownames(mat2)=1:3
mat2
```

create mat2D as data frame of mat2

```{r}
mat2D=as.data.frame(mat2)
mat2D
```

### Binding data sets

```{r}
colnames(mat2D)=colnames(mat1D) #Make column names same
rbind(mat1D,mat2D)
```

# DataFrame

2-dimension data structure consists of different types of data

```{r}
no=1:4
name=c("Apple", "Banana", "Peach", "Strawberry")
price=c(500,400,300,200)
quantity=c(5,2,4,7)
```

```{r}
fruit=data.frame(no=no,item=name,Price=price, Stock=quantity)
fruit
```

```{r}
str(fruit)
```

```{r}
fruit[,c(2,4)]
```

```{r}
fruit[,-1]
```

```{r}
fruit[-c(2,4),-1]
```

```{r}
fruit$StockPrice=fruit$Price*fruit$Stock
fruit
```

```{r}
gold=read.csv("gold.csv",header=T)
gold
```

```{r}
fruit$Site=c("Kor","Tha","Kor","Chn")
fruit
```

# Loading external file

```{r}
txtEmp=read.table("emp.txt", header=T, sep=",")
txtEmp
```

```{r}
str(txtEmp)
```

```{r}
csvEmp=read.csv("emp2.csv", header=F)
csvEmp
```

```{r}
colnames(csvEmp)=c("사번", "이름", "급여")
csvEmp
```

```{r}
ncol(csvEmp)
nrow(csvEmp)
```

```{r}
df=data.frame(x=1:5,y=seq(2,10,2),z=c("a","b","c","d","e"))
df
```

```{r}
sum(df$x)
sum(df$y)
```

# Function calls function

```{r}
apply(df[,-3], 2, sum)
apply(df[,-3], 2, mean)
apply(df[,-3], 2, sd)
```

```{r}
df$sum=apply(df[,-3], 1, sum)
df
```

subset(like SELECT in SQL)

```{r}
nrow(subset(df,x>=3&y>=8))
```

```{r}
apply(subset(df,x>=3&y>=8)[,1:2],2,sum)
```

# Merging multiple data frame

Like join in SQL

-   Key is same and quantity of data is same

```{r}
height=data.frame(id=1:3, height=c(160,170,180))
weight=data.frame(id=c(2,1,3),weight=c(70,60,80))
height;weight
```

Automatically matches key with merge()

```{r}
merge(height,weight)
```

-   Different key column, different quantity of data

```{r}
height=data.frame(id=1:3,height=c(160,170,180))
weight=data.frame(ad=c(2,1),weight=c(70,60))
height;weight
```

```{r}
merge(height,weight)
```

Like inner join in SQL

```{r}
merge(x=height, y=weight, by.x="id", by.y="ad")
```

Right outer join in SQL

```{r}
merge(x=height, y=weight, by.x="id", by.y="ad",all.x=T)
```

# List

```{r}
member=list(name="Carlson", address="Seoul", tel="010-1111-2222", pay=300);member
```

```{r}
c(member$name, member$address)
```

# String

Cut, replace, extract string in data collected

## stringr package

-   str_length(): length of string

-   str_c(): connect string

-   str_sub(): substring

-   str_split(): split string according to splitter

-   str_replace(): replace original string to new one

-   str_extract(), str_extract_all(): extract first(all) of selected string pattern

-   str_locate(), str_locate_all(): first(all) location of selected string patttern

```{r}
library(stringr) #import in Java
```

```{r}
string="magnuscarlson2859dingliren2811이준혁2388"
string
```

```{r}
str_extract_all(string,"[a-z]{3}")
```

```{r}
str_extract_all(string,"[a-z]{3,5}")
str_extract_all(string,"[a-z]{3,}") #list
```

```{r}
aaa=str_extract_all(string,"[a-z]{3,}")
bbb=unlist(aaa) #character
bbb
```

```{r}
ccc=str_extract_all(string,"[0-9]{1,}");ccc
ddd=unlist(ccc)
ddd;ddd[1]
```

```{r}
eee=str_extract_all(string,"[가-힣]{1,}");eee
```

```{r}
str_extract_all(string,"[^a-z]{1,}") #pattern except alphabet
```

```{r}
str_length(string)
str_locate_all(string,"magnus")
class(str_locate_all(string,"magnus"))
```

```{r}
str_sub(string,1,32)
str_sub(string,1, str_locate(string,"ding")[1]-1)
```

```{r}
str_to_upper(string)
str_to_lower(string)
```

```{r}
string=str_replace(string,"이준혁2388", "leejunhyeok2388")
string
```

```{r}
string2="magnuscarlson2859,dingliren2811,이준혁2388"
unlist(str_split(string2,","))
```

```{r}
price="12,345,678"
as.numeric(str_replace_all(price,",",""))
as.numeric(str_remove_all(price,","))
```

```{r}
str_c(price,sep="")
```

#### Replace NA

```{r}
temp=c(1,2,NA,4)
sum(is.na(temp))
str_replace_na(temp,median(temp,na.rm=T))
```
